---
title: "QPCR"
author: "QUINN LE"
date: "DATE"
output: html_document
---
#  Load packages
```{r setup, include=FALSE}
rm(list = ls())
library(tidyverse)
library(ggplot2)
library(magrittr)
library(readxl)
library(data.table)


library(dplyr)
`%!in%` = Negate(`%in%`)

```
# Please load data and name as 'dlp' for dloop data and 'alb' for albumin data
# Quality control: 
## Check water has no signal
## Drop samples that have no signal for mtDNA-CN
```{r}
# INPUT CORRESPONDING CSV PATH
dlp <- read.csv("/Users/quinnmle/Documents/Castellani Lab/Mito_2026-03-03/dlp.csv")

alb <- read.csv("/Users/quinnmle/Documents/Castellani Lab/Mito_2026-03-03/alb_adj.csv")

removed_samples = as.data.frame(matrix(nrow = 1, ncol = 3))
colnames(removed_samples) = c('Sample.Name', 'Well', 'Reason')
dlp$Dlp_CT = dlp$Cт
alb$Alb_CT = alb$Cт
df = merge(dlp, alb, by = c('Well', 'Sample.Name')) 
df$Dlp_CT <- as.numeric(as.character(df$Dlp_CT))
df$Alb_CT <- as.numeric(as.character(df$Alb_CT))
no_mt = subset(df, Dlp_CT > 26)
#no_mt$Reason = 'D-loop CT > 26'
#removed_samples = rbind(removed_samples, dplyr::select(no_mt, `Sample.Name`, `Well`, Reason))
#subset(no_mt, `Sample.Name` == 'water') # check that water is here.
df = subset(df, Dlp_CT <= 26)
```
## Calculate deltaCT
```{r}
df$deltaCT = df$Alb_CT - df$Dlp_CT
```
## Check for pipetting effects
```{r}
mydata_deltaCT = df
colnames(mydata_deltaCT) = make.names(colnames(mydata_deltaCT))
# Repurposed from Foram:
rank.deltaCT = data.frame(matrix(NA, ncol = 6))
names(rank.deltaCT) = c("Sample.Name", "Min.CT", "Mid.CT", 
                        "Max.CT", "CT.Dlp.diff", "CT.Alb.diff")
samplelist = levels(as.factor(mydata_deltaCT$Sample.Name))
for (i in 1:length(samplelist)) {
  test = subset(mydata_deltaCT, na.omit(Sample.Name == 
                                          samplelist[i]))
  test$Well = as.numeric(gsub("[A-Z]", "", test$Well))
  test.ordered = test[order(test$Well), ]
  test.ordered$deltaCT = as.numeric(test.ordered$deltaCT)
  if (nrow(test) == 3) {
    first = which(test.ordered$deltaCT == min(test.ordered$deltaCT), 
                  arr.ind = T)
    second = which(test.ordered$deltaCT == median(test.ordered$deltaCT), 
                   arr.ind = T)
    third = which(test.ordered$deltaCT == max(test.ordered$deltaCT), 
                  arr.ind = T)
    sample = test.ordered$Sample.Name[1]
    ct.dlp.diff = test.ordered$Dlp_CT[1] - test.ordered$Dlp_CT[3]
    ct.alb.diff = test.ordered$Alb_CT[1] - test.ordered$Alb_CT[3]
    rank.deltaCT[i, ] = c(test$Sample.Name[1], first, 
                          second, third, ct.dlp.diff, ct.alb.diff)
  }
}
rank.deltaCT = na.omit(rank.deltaCT)
rank.deltaCT$CT.Dlp.diff = as.numeric(rank.deltaCT$CT.Dlp.diff)
rank.deltaCT$CT.Alb.diff = as.numeric(rank.deltaCT$CT.Alb.diff)
Min = table(rank.deltaCT$Min.CT)
Mid = table(rank.deltaCT$Mid.CT)
Max = table(rank.deltaCT$Max.CT)
#t = rbind(Min, Mid, Max)
# Average CT difference between first and third replicate
mean(rank.deltaCT$CT.Dlp.diff) 
mean(rank.deltaCT$CT.Alb.diff)
# Is there evidence of a pipetting order effect?
```
# Assign + adjust(?) for pipetting order
```{r}
data.deltaCT = mydata_deltaCT
data.deltaCT.PO=data.frame(matrix(NA, ncol=6))
colnames(data.deltaCT.PO)=c("Sample.Name","deltaCT","Well","PO", "Alb_CT", "Dlp_CT")
samplelist=levels(as.factor(data.deltaCT$`Sample.Name`))
for (i in 1:length(samplelist)){
  test=subset(data.deltaCT, Sample.Name==samplelist[i])
  test$Well.number=as.numeric(gsub("[A-Z]","",test$Well))
  test$Well.letter=gsub("[0-9]","",test$Well)
  test.ordered=test[order(test$Well.letter, test$Well.number),]
  test.ordered$deltaCT=as.numeric(test.ordered$deltaCT)
  if (nrow(test.ordered)>1){
    test.ordered$PO=1:nrow(test.ordered)
    test.ordered = dplyr::select(test.ordered, "Sample.Name","deltaCT","Well","PO", "Alb_CT", "Dlp_CT")
    data.deltaCT.PO=rbind(data.deltaCT.PO, test.ordered)
  }
}
data.deltaCT.PO <- na.omit(data.deltaCT.PO)
# Visualize pipetting effect
ggplot(data.deltaCT.PO, aes(as.factor(PO), deltaCT)) + geom_boxplot() + xlab('Pipetting order')
# Adjust for PO?
lm(deltaCT ~ PO, data = data.deltaCT.PO) %>% summary
# Not significant, not going to adjust.
# data.deltaCT.PO$deltaCT.POadjust = resid(lm(deltaCT ~ PO, data = data.deltaCT.PO))
df = data.deltaCT.PO
```
## Drop samples that have too much variation between replicates
```{r}
colnames(removed_samples) = make.names(colnames(removed_samples))
library(raster)
# Get coefficients of variation for each sample
df %>% group_by(Sample.Name) %>% summarise(cv= cv(deltaCT)) -> coefs
# Plot density
ggplot(coefs, aes(cv)) + geom_density()
# For samples with CVs greater than 15%, drop the biggest outlier and recheck CV
high = subset(coefs, cv > 15)
for(i in 1:nrow(high))
{
  test = subset(df, Sample.Name == high$Sample.Name[i])
  
  if(nrow(test) > 2){
    
    # testing standard deviation of different combos
    # Makes new variables named by Wellpositions
    assign(test$Well[1],sd(test$deltaCT[2:3])) 
    assign(test$Well[2],sd(test$deltaCT[c(1,3)]))
    assign(test$Well[3],sd(test$deltaCT[1:2]))
    bad.sample=min(as.numeric(mget(test$Well)))
    bad.sample2=test[which(bad.sample==mget(test$Well)),c("Sample.Name","Well")]
    bad.sample2$Reason = 'CV too high'
    
    # Add to removed samples
    removed_samples = plyr::rbind.fill(removed_samples, bad.sample2)
    
    # Drop the most extreme
    df %>% filter(Well != bad.sample2$Well) -> df
  }
}
```
## Drop anything that still has a high CV
```{r}
df %>% group_by(Sample.Name) %>% summarise(cv= cv(deltaCT)) -> coefs
bad.samps = subset(coefs, cv > 15)
bad = subset(df, Sample.Name %in% bad.samps$Sample.Name)
#bad$Reason = 'CV still too high after filtering'
#removed_samples = rbind(removed_samples, bad)
df = subset(df, Sample.Name %!in% bad.samps$Sample.Name)
```
## Remove samples with only one replicate left
```{r}
df %>% group_by(Sample.Name) %>% summarise(count= n()) -> number
number$Sample.Name[which(number$count == 1)]
oneleft = subset(df, Sample.Name %in% number$Sample.Name[which(number$count == 1)])
#oneleft$Reason = 'Only one replicate left'
removed_samples = rbind(removed_samples, oneleft)
df = subset(df, Sample.Name %!in% number$Sample.Name[which(number$count == 1)])
```

# QC
```{r}


data_for_filtering <- df
data_for_filtering$Sample <- data_for_filtering$Sample.Name
print("n to move forward for filtering, including controls:")
nrow(data_for_filtering)
print("n to move forward for filtering, no controls:")
subset(data_for_filtering, Sample != "Water") %>% nrow()
data_for_filtering <- subset(data_for_filtering, Sample != "Water")

```

## Samples with NA DLP CT or ALB CT
```{r}
DLP_ALB_CT_NA_sample <- subset(data_for_filtering, is.na(Dlp_CT) | is.na(Alb_CT))
print("n samples with NA DLP_CT or ALB_CT")
nrow(DLP_ALB_CT_NA_sample)
```

## Samples with DLP CT greater than 26
```{r}
DLP_over26_sample <- subset(data_for_filtering, Dlp_CT > 26)
print("n samples with DLP_CT > 26")
nrow(DLP_over26_sample)
```

## Samples with ALB CT more than 3 SD in any direction
```{r}
highALB <- mean(data_for_filtering$Alb_CT) + (2* (sd(data_for_filtering$Alb_CT)))
lowALB <- mean(data_for_filtering$Alb_CT) - (2* (sd(data_for_filtering$Alb_CT)))

highALB_sample <- subset(data_for_filtering, Alb_CT > highALB)
print("highALB samples")
nrow(highALB_sample)

lowALB_sample <- subset(data_for_filtering, Alb_CT < lowALB)
print("lowALB samples")
nrow(lowALB_sample)

# lowALB_sample$qc_assessment <- c('ALB CT < 3 SD from the mean')
# highALB_sample$qc_assessment <- c('ALB CT > 3 SD from the mean')
```

## Samples with deltaCT more than 3 SD in any direction
```{r}
highDCT <- mean(data_for_filtering$deltaCT) + (3*(sd(data_for_filtering$deltaCT)))
lowDCT <- mean(data_for_filtering$deltaCT) - (3*(sd(data_for_filtering$deltaCT)))


highDCT_sample <- subset(data_for_filtering, deltaCT > highDCT)
print("high deltaCT samples")
nrow(highDCT_sample)

lowDCT_sample <- subset(data_for_filtering, deltaCT < lowDCT)
print("low deltaCT samples")
nrow(lowDCT_sample)

#lowDCT_sample$qc_assessment <- c('deltaCT < 3 SD from the mean')
#highDCT_sample$qc_assessment <- c('deltaCT > 3 SD from the mean')
```

## Removing samples
```{r}
data_for_filtering <- subset(data_for_filtering, Dlp_CT <= 26)
print("n samples after removing >26 DLP CT")
subset(data_for_filtering, Dlp_CT <= 26) %>% nrow()
after_df <- subset(data_for_filtering, Dlp_CT <= 26)

data_for_filtering <- subset(data_for_filtering, Alb_CT > lowALB & Alb_CT < highALB)
print("n samples after removing extreme ALB")
subset(data_for_filtering, Alb_CT > lowALB & Alb_CT < highALB) %>% nrow()
after_df <- subset(after_df, Alb_CT > lowALB & Alb_CT < highALB)

data_for_filtering <- subset(data_for_filtering, deltaCT > lowDCT & deltaCT < highDCT)
print("n samples after removing extreme delta CT")
subset(data_for_filtering, deltaCT > lowDCT & deltaCT < highDCT) %>% nrow()
after_df <- subset(after_df, deltaCT > lowDCT & deltaCT < highDCT)
```
## Fold Change
```{r}
after_df$Sample <- as.factor(after_df$Sample)


order(after_df$Sample)
after_df

# WT_NC_dCt_mean <- mean(after_df$deltaCT[after_df$Sample.Name == "NC3.Sham.1"])
# KO11_NC_dCt_mean <- mean(after_df$deltaCT[after_df$Sample.Name == "KO5.Sham.1"])
# KO11_A_dCt_mean <- mean(after_df$deltaCT[after_df$Sample.Name == "KO5.A.1"])
# KO11_B_dCt_mean <- mean(after_df$deltaCT[after_df$Sample.Name == "KO5.B.1"])
# KO11_C_dCt_mean <- mean(after_df$deltaCT[after_df$Sample.Name == "KO5.C.1"])
# KO11_D_dCt_mean <- mean(after_df$deltaCT[after_df$Sample.Name == "KO5.D.1"])

WT_NC_dCt_mean <- mean(after_df$deltaCT[after_df$Sample.Name == "WT_SHAM"])
KO11_NC_dCt_mean <- mean(after_df$deltaCT[after_df$Sample.Name == "KO_SHAM"])
KO11_A_dCt_mean <- mean(after_df$deltaCT[after_df$Sample.Name == "KO_A"])
KO11_B_dCt_mean <- mean(after_df$deltaCT[after_df$Sample.Name == "KO_B"])
KO11_C_dCt_mean <- mean(after_df$deltaCT[after_df$Sample.Name == "KO_C"])
KO11_D_dCt_mean <- mean(after_df$deltaCT[after_df$Sample.Name == "KO_D"])


fold_change <- c(2^(KO11_A_dCt_mean-KO11_NC_dCt_mean), 
                 2^(KO11_B_dCt_mean-KO11_NC_dCt_mean), 
                 2^(KO11_C_dCt_mean-KO11_NC_dCt_mean),
                 2^(KO11_D_dCt_mean-KO11_NC_dCt_mean))

fold_change

library(dplyr)
library(ggplot2)

# ---- Step 1: Define groups ----
plot_data <- data %>%
  mutate(
    group = case_when(
      type == "KO_A" ~ "KO_A",
      type == "KO_B" ~ "KO_B",
      type == "KO_C" ~ "KO_C",
      type == "KO_D" ~ "KO_D",
      type == "KO_SHAM" ~ "KO_SHAM",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(group))

# ---- Step 2: Get reference mean (KO_SHAM) ----
ref_mean <- plot_data %>%
  filter(group == "KO_SHAM") %>%
  summarise(mean_ref = mean(deltaCT, na.rm = TRUE)) %>%
  pull(mean_ref)

# ---- Step 3: Summarise ALL groups ----
fc_df <- plot_data %>%
  group_by(group) %>%
  summarise(
    n = sum(!is.na(deltaCT)),
    mean_dCt = mean(deltaCT, na.rm = TRUE),
    sd_dCt = sd(deltaCT, na.rm = TRUE),
    se_dCt = sd_dCt / sqrt(n),
    tcrit = qt(0.975, df = n - 1),  # 90% CI
    .groups = "drop"
  ) %>%
  mutate(
    ddct = mean_dCt - ref_mean,
    ddct_lower = ddct - tcrit * se_dCt,
    ddct_upper = ddct + tcrit * se_dCt,

    # YOUR definition
    fold_change = 2^(ddct),
    fc_lower = 2^(ddct_lower),
    fc_upper = 2^(ddct_upper)
  ) %>%
  filter(group != "KO_SHAM") %>%  # remove reference from plot if you want
  mutate(
    group = factor(group, levels = c("KO_A", "KO_B", "KO_C", "KO_D"))
  )

# ---- Step 4: Plot ----
p <- ggplot(fc_df, aes(x = group, y = fold_change)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  
  geom_point(size = 3) +
  
  geom_errorbar(
    aes(ymin = fc_lower, ymax = fc_upper),
    width = 0.12
  ) +
  
  scale_y_continuous(
    breaks = c(0, 1, 2, 4, 6, 8, 10)
  ) +
  
  labs(
    title = "Fold Change vs KO Sham",
    x = NULL,
    y = "Fold Change"
  ) +
  
  theme_classic()

print(p)

# ---- Step 5: Save ----
ggsave(
  "fold_change_CI.png",
  plot = p,
  width = 6,
  height = 4,
  dpi = 300
)



```




## Replot after QC
```{r}
after_df <- data_for_filtering

# plot_grid(ggplot(after_df, aes(x = Alb_CT %>% as.numeric())) +
#             geom_density(alpha = 0.5) + xlab('ALB CT') + ylab('Density') +
#             geom_rug() + theme_bw(),
#           ggplot(after_df, aes(x = Dlp_CT %>% as.numeric())) +
#             geom_density(alpha = 0.5) + xlab('DLP CT') + ylab('Density') +
#             geom_rug() + theme_bw(),
#           ggplot(after_df, aes(x = deltaCT %>% as.numeric())) +
#             geom_density(alpha = 0.5) + xlab('Delta CT') + ylab('Density') +
#             geom_rug() + theme_bw(),
#           nrow = 1, ncol = 3)
ggplot(after_df, aes(x = Alb_CT %>% as.numeric())) +
            geom_density(alpha = 0.5) + xlab('ALB CT') + ylab('Density') +
            geom_rug() + theme_bw()
          ggplot(after_df, aes(x = Dlp_CT %>% as.numeric())) +
            geom_density(alpha = 0.5) + xlab('DLP CT') + ylab('Density') +
            geom_rug() + theme_bw()
          ggplot(after_df, aes(x = deltaCT %>% as.numeric())) +
            geom_density(alpha = 0.5) + xlab('Delta CT') + ylab('Density') +
            geom_rug() + theme_bw()
```

## Summary of low quality samples to be removed
```{r}
print("Samples with ALB CT > 3 SD in any direction")
lowALB_sample$Sample 
highALB_sample$Sample

print("Samples with DLP CT > 26")
DLP_over26_sample$Sample

print("Samples with delta CT > 3 SD in any direction")
lowDCT_sample$Sample
highDCT_sample$Sample
```

```{r}
df <- after_df
#df <- subset(df, df$Sample.Name != "KO11_B")
setwd("/Users/quinnmle/Documents/Castellani Lab/Mito_2026-03-18")
write.csv(df, "qpcr_results.csv", quote=F)

df <- subset(df, df$Sample.Name != "hgdna")

df2 <- df[,c("Sample.Name", "deltaCT", "Alb_CT","Dlp_CT")]
data.deltaCT.PO.clean <- df2
 meanlist=levels(as.factor(data.deltaCT.PO.clean$Sample.Name))
  data.means=data.frame(matrix(NA, ncol=4))
  names(data.means)=c("Sample","deltaCT.mean","CT.Alb", "CT.Dlp")
  for (i in 1:length(meanlist)){
    sample.test=meanlist[i]
    if (nrow(na.omit(subset(data.deltaCT.PO.clean, Sample.Name==sample.test)))==1){
      mean.delta=mean.adjusted=mean.Dlp=mean.Alb=NA
    } else {
      mean.delta=mean(as.numeric(na.omit(subset(data.deltaCT.PO.clean, Sample.Name==sample.test)$deltaCT)))
      mean.Dlp=mean(as.numeric(na.omit(subset(data.deltaCT.PO.clean, Sample.Name==sample.test)$Dlp_CT)))
      mean.Alb=mean(as.numeric(na.omit(subset(data.deltaCT.PO.clean, Sample.Name==sample.test)$Alb_CT)))
       sd.test=sd(as.numeric(na.omit(subset(data.deltaCT, Sample.Name==sample.test)$deltaCT)))
    }
    data.means[i,]=c(sample.test,mean.delta,mean.Alb, mean.Dlp)
    data.means$deltaCT.mean <- as.numeric(data.means$deltaCT.mean)
  }
 # merged.means <- rbind(merged.means, data.means)
#Data Analysis
 # merged.clean.replicates <- merged.clean.replicates[!is.na(merged.clean.replicates$Sample),]
#  merged.clean.replicates$Sample <- as.factor(merged.clean.replicates$Sample)
  #merged.means <- merged.means[!is.na(merged.means$Sample),]
#  data.means$Sample.Name <- as.factor(data.means$Sample.Name)

  # EDIT LABELS AND RANGES!
forPlot <- df2[order(df2$Sample),]
#forPlot$type <- c("NC")
#forPlot$type[1:3] <- c("TFAM KO11")
#forPlot$type[4:6] <- c("TFAM KO5")
#forPlot$type[7:8] <- c("TFAM KO6")
#forPlot$type[9:11] <- c("TFAM NC1")
#forPlot$type[12:14] <- c("TFAM NC2")
#forPlot$type[15:17] <- c("TFAM NC")

data <- forPlot
for (i in 1:nrow(data)){
  data$type[i]=data$Sample.Name[i]
}
#data <- forPlot
library(ggplot2)

# ARRANGE THE LABELS ACCORDING TO THE ORDER YOU WANT TO SEE THEM IN"
data$type <- factor(data$type, levels=c("WT_SHAM","KO_SHAM", "KO_C", "KO_B", "KO_A", "KO_D"))

#setwd("/Users/Christina/Desktop/MonochromePlates/")

data$Dlp_CT <- as.numeric(data$Dlp_CT)
data$Alb_CT <- as.numeric(data$Alb_CT)
minC <- min(data$deltaCT)
minD <- min(data$Dlp_CT)
minA <- min(data$Alb_CT)

# NAME YOUR PDF, GGTITLE SHOULD BE YOUR ORIGINAL CSV
pdf(file = "deltaCt_adj.pdf", width = 10)
ggplot(data) + geom_boxplot(aes(col=type, x=type, y=deltaCT)) +  ggtitle("deltaCt_adj") + ylab("deltaCt") + theme(legend.title=element_blank(), axis.title = element_text(face="bold")) + geom_point(aes(col=type, x=type, y=deltaCT), position=position_dodge(0.75)) +  theme(axis.text.x=element_text(angle = 45, hjust = 1), plot.margin = margin(1, 0.2, 1.5, 2, "cm"), text = element_text(size=10), axis.title.x=element_blank(), legend.title=element_blank()) + ylim(minC-0.2, minC+5.5)


dev.off()


# NAME YOUR PDF, GGTITLE SHOULD BE YOUR ORIGINAL DLP CSV
pdf(file = "dlp.pdf", width = 10)
ggplot(data) + geom_boxplot(aes(col=type, x=type, y=Dlp_CT)) +  ggtitle("Dlp") + ylab("Dlp_CT") + theme(legend.title=element_blank(), axis.title = element_text(face="bold")) + geom_point(aes(col=type, x=type, y=Dlp_CT), position=position_dodge(0.75)) +  theme(axis.text.x=element_text(angle = 45, hjust = 1), plot.margin = margin(1, 0.2, 1.5, 2, "cm"), text = element_text(size=10), axis.title.x=element_blank(), legend.title=element_blank()) + ylim(minD-0.2,minD+5.5)
dev.off()

# NAME YOUR PDF, GGTITLE SHOULD BE YOUR ORIGINAL ALB CSV
pdf(file = "alb_adj.pdf", width = 10)
ggplot(data) + geom_boxplot(aes(col=type, x=type, y=Alb_CT)) +  ggtitle("Alb_adj") + ylab("Alb_CT") + theme(legend.title=element_blank(), axis.title = element_text(face="bold")) + geom_point(aes(col=type, x=type, y=Alb_CT), position=position_dodge(0.75)) +  theme(axis.text.x=element_text(angle = 45, hjust = 1), plot.margin = margin(1, 0.2, 1.5, 2, "cm"), text = element_text(size=10), axis.title.x=element_blank(), legend.title=element_blank()) + ylim(minA-0.2, minA+5.5)
dev.off()

```
```{r}
 # Make sure types are set and deltaCT is numeric
data$type <- droplevels(as.factor(data$Sample.Name))
data$deltaCT <- as.numeric(data$deltaCT)

# Pairwise t-tests (Welch by default); adjust p-values
pw <- pairwise.t.test(
  x = data$deltaCT,
  g = data$type,
  p.adjust.method = "bonferroni"   # or "bonferroni"
)


t.test(deltaCT ~ type, data = subset(data, type %in% c("KO_SHAM", "KO_A")))
t.test(deltaCT ~ type, data = subset(data, type %in% c("KO_SHAM", "KO_B")))
t.test(deltaCT ~ type, data = subset(data, type %in% c("KO_SHAM", "KO_C")))
# t.test(deltaCT ~ type, data = subset(data, type %in% c("KO_SHAM", "KO_D")))
t.test(deltaCT ~ type, data = subset(data, type %in% c("KO_SHAM", "WT_SHAM")))
t.test(deltaCT ~ type, data = subset(data, type %in% c("WT_SHAM", "KO_A")))
t.test(deltaCT ~ type, data = subset(data, type %in% c("WT_SHAM", "KO_B")))
t.test(deltaCT ~ type, data = subset(data, type %in% c("WT_SHAM", "KO_C")))
# t.test(deltaCT ~ type, data = subset(data, type %in% c("WT_SHAM", "KO_D")))



pw
# pw$p.value

```



```{r}


library(ggplot2)
library(ggsignif)

data$type <- factor(data$type, levels = c("WT_SHAM", "KO_SHAM", "KO_A", "KO_B", "KO_C", "KO_D"))

ymax <- max(data$deltaCT, na.rm = TRUE)
ymin <- min(data$deltaCT, na.rm = TRUE)

p <- ggplot(data, aes(x = type, y = deltaCT, color = type)) +
  geom_boxplot(outlier.shape = NA, width = 0.6) +
  geom_jitter(width = 0.12, size = 2) +
  geom_signif(
    comparisons = list(
      c("KO_SHAM", "KO_A"),
      c("KO_SHAM", "KO_B"),
      c("KO_SHAM", "KO_C"),
      c("KO_SHAM", "KO_D"),
      c("WT_SHAM", "KO_SHAM")
    ),
    y_position = c(ymax + 0.3, ymax + 0.8, ymax + 1.3, ymax + 1.8),
    tip_length = 0.02,
    map_signif_level = TRUE,
    test = t.test,
    color = "black",
    textsize = 4
  ) +
  labs(title = "deltaCt_adj", x = NULL, y = "deltaCt") +
  theme_classic(base_size = 12) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(face = "bold")
  ) +
  coord_cartesian(ylim = c(ymin - 0.2, ymax + 2.3))

print(p)

ggsave("deltaCt_adj.png", path = "/Users/quinnmle/Documents/Castellani Lab/Mito_2026-03-18", plot = p, width = 10, height = 6)



```



```{r}
library(dplyr)
library(ggplot2)
fc_summary_all <- plot_data %>%
  group_by(plot_group) %>%
  summarise(
    n = sum(!is.na(deltaCT)),
    mean_deltaCT = mean(deltaCT, na.rm = TRUE),
    sd_deltaCT = sd(deltaCT, na.rm = TRUE),
    se_deltaCT = sd_deltaCT / sqrt(n),
    tcrit = ifelse(n > 1, qt(0.95, df = n - 1), NA_real_),  # 90% CI
    .groups = "drop"
  ) %>%
  mutate(
    ddct = mean_deltaCT - ref_mean,
    ddct_lower = ifelse(n > 1, ddct - tcrit * se_deltaCT, NA_real_),
    ddct_upper = ifelse(n > 1, ddct + tcrit * se_deltaCT, NA_real_),
    fold_change = 2^(-ddct),
    fc_lower = ifelse(n > 1, 2^(-ddct_upper), NA_real_),
    fc_upper = ifelse(n > 1, 2^(-ddct_lower), NA_real_)
  )
fc_summary_all$plot_group <- factor(
  fc_summary_all$plot_group,
  levels = c(
    "Wild-Type Sham\nTransplant",
    "TFAM-KO Sham\nTransplant",
    "KO_A",
    "KO_B",
    "KO_C",
    "KO_D"
  )
)
ggplot(fc_summary_all, aes(x = plot_group, y = fold_change)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  
  geom_point(size = 3) +
  
  geom_errorbar(
    aes(ymin = fc_lower, ymax = fc_upper),
    width = 0.12,
    na.rm = TRUE
  ) 
  
  labs(
    title = "Fold Change vs TFAM-KO Sham Transplant",
    x = NULL,
    y = "Fold Change (2^-ddCt, 90% CI)"
  ) +
  
  theme_classic()
```
