---
title: "RNA-Seq Thesis"
author: "Quinn Le"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set-Up
```{r, echo=FALSE}


library(DESeq2)
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(ggplot2)
library(ggrepel)
library(AnnotationDbi)
library(enrichplot)


setwd("/Users/quinnmle/Documents/Documents/Thesis Analysis")

```


## Input File Set-Up
```{r}
counts_file_path <- "counts.txt"
sample_info_file_path <- "sample_metadata.csv"
```


# Importing Count Matrix
```{r}
counts <- read.table(counts_file_path, sep = "\t", header = TRUE)
rownames(counts) <- counts$Geneid 
counts <- counts[,-(1:6)]
counts <- as.matrix(counts)
```

## Importing Sample Information
```{r}
sample_info <- read.csv(sample_info_file_path, header = TRUE)
rownames(sample_info) <- sample_info$sample
sample_info$condition <- as.factor(sample_info$condition)

print("Comparison Groups:")
levels(sample_info$condition)
```

# Creating DESeq2 Object
```{r}
# Only works for this
colnames(counts) <- rownames(sample_info)


dds <- DESeqDataSetFromMatrix(
  countData = counts,
  colData = sample_info,
  design = ~ condition
)

print("Number of genes before filtering:")
nrow(dds)

dds <- dds[rowSums(counts(dds)) >= 10, ]

print("Number of genes after removing those with less than 10 counts:")
nrow(dds)

dds <- DESeq(dds, useT = TRUE)
```

```{r}

all_genes_high <- as.data.frame(results(dds, contrast = c("condition", "mito_high", "control")))
all_genes_high$gene <- rownames(all_genes_high)

all_genes_low <- as.data.frame(results(dds, contrast = c("condition", "mito_low", "control")))
all_genes_low$gene <- rownames(all_genes_low)

all_genes_hl <- as.data.frame(results(dds, contrast = c("condition", "mito_high", "mito_low")))
all_genes_hl$gene <- rownames(all_genes_hl)

```



Exploring number of signficant genes
```{r}
cat("Number of genes with padj < 0.05","\n")
cat(paste("High vs. ctrl:", 
            nrow(subset(all_genes_high, 
                        !is.na(padj) & padj < 0.05))),"\n")
cat(paste("Low vs. ctrl:", 
            nrow(subset(all_genes_low, 
                        !is.na(padj) & padj < 0.05))),"\n")
cat(paste("High vs. Low:", 
            nrow(subset(all_genes_hl, 
                        !is.na(padj) & padj < 0.05))),"\n\n")

cat("Number of genes with padj < 0.01","\n")
cat(paste("High vs. ctrl:", 
            nrow(subset(all_genes_high, 
                        !is.na(padj) & padj < 0.01))),"\n")
cat(paste("Low vs. ctrl:", 
            nrow(subset(all_genes_low, 
                        !is.na(padj) & padj < 0.01))),"\n")
cat(paste("High vs. Low:", 
            nrow(subset(all_genes_hl,
                        !is.na(padj) & padj < 0.01))),"\n\n")

cat("Number of genes with padj < 0.05 and log2FC > 1","\n")
cat(paste("High vs. ctrl:", 
            nrow(subset(all_genes_high, 
                        !is.na(padj) & padj < 0.05 & abs(log2FoldChange) > 1))),"\n")
cat(paste("Low vs. ctrl:", 
            nrow(subset(all_genes_low, 
                        !is.na(padj) & padj < 0.05 & abs(log2FoldChange) > 1))),"\n")
cat(paste("High vs. Low:", 
            nrow(subset(all_genes_hl, 
                        !is.na(padj) & padj < 0.05 & abs(log2FoldChange) > 1))),"\n\n")

cat("Number of genes with padj < 0.01 and log2FC > 1","\n")
cat(paste("High vs. ctrl:", 
            nrow(subset(all_genes_high, 
                        !is.na(padj) & padj < 0.01 & abs(log2FoldChange) > 1))),"\n")
cat(paste("Low vs. ctrl:", 
            nrow(subset(all_genes_low, 
                        !is.na(padj) & padj < 0.01 & abs(log2FoldChange) > 1))),"\n")
cat(paste("High vs. Low:", 
            nrow(subset(all_genes_hl, 
                        !is.na(padj) & padj < 0.01 & abs(log2FoldChange) > 1))),"\n\n")

cat("Number of genes with padj < 0.05 and log2FC > 2","\n")
cat(paste("High vs. ctrl:", 
            nrow(subset(all_genes_high, 
                        !is.na(padj) & padj < 0.05 & abs(log2FoldChange) > 2))),"\n")
cat(paste("Low vs. ctrl:", 
            nrow(subset(all_genes_low, 
                        !is.na(padj) & padj < 0.05 & abs(log2FoldChange) > 2))),"\n")
cat(paste("High vs. Low:", 
            nrow(subset(all_genes_hl, 
                        !is.na(padj) & padj < 0.05 & abs(log2FoldChange) > 2))),"\n")

cat("Number of genes with padj < 0.01 and log2FC > 2","\n")
cat(paste("High vs. ctrl:", 
            nrow(subset(all_genes_high, 
                        !is.na(padj) & padj < 0.01 & abs(log2FoldChange) > 2))),"\n")
cat(paste("Low vs. ctrl:", 
            nrow(subset(all_genes_low, 
                        !is.na(padj) & padj < 0.01 & abs(log2FoldChange) > 2))),"\n")
cat(paste("High vs. Low:", 
            nrow(subset(all_genes_hl, 
                        !is.na(padj) & padj < 0.01 & abs(log2FoldChange) > 2))),"\n\n")
```

Subsetting significant genes
```{r}
# Subsetting significant genes
sig_genes_high <- subset(all_genes_high, !is.na(padj) & padj < 0.05 & abs(log2FoldChange) > 2)
sig_genes_low <- subset(all_genes_low, !is.na(padj) & padj < 0.05 & abs(log2FoldChange) > 2)
sig_genes_hl <- subset(all_genes_hl, !is.na(padj) & padj < 0.05 & abs(log2FoldChange) > 2)
```

Removing Ensembl version suffix if present
```{r}
sig_genes_high$gene <- sub("\\..*$", "", sig_genes_high$gene)
sig_genes_low$gene <- sub("\\..*$", "", sig_genes_low$gene)
sig_genes_hl$gene <- sub("\\..*$", "", sig_genes_hl$gene)
```

Function for printing number of upregulated and downregulated genes
```{r}
calculate_up_down <- function(res_obj, sig_cutoff = 0.05, lfc_cutoff=2) {
  up <- subset(res_obj, !is.na(padj) & padj < sig_cutoff & log2FoldChange > lfc_cutoff)
  down <- subset(res_obj, !is.na(padj) & padj < sig_cutoff & log2FoldChange < -lfc_cutoff)
  cat(deparse(substitute(res_obj)),"\n")
  cat(paste("Number of significant genes:   ", nrow(res_obj)),"\n")
  cat(paste("Number of upregulated genes:   ", nrow(up)),"\n")
  cat(paste("Number of downregulated genes: ", nrow(down)),"\n\n")
  write.csv(up, paste0(deparse(substitute(res_obj)), "_upregulated.csv"))
  write.csv(down, paste0(deparse(substitute(res_obj)), "_downregulated.csv"))
}
```

Running function
```{r}
calculate_up_down(sig_genes_high)
calculate_up_down(sig_genes_low)
calculate_up_down(sig_genes_hl)
```


# PCA
```{r}

# vsd <- vst(dds, blind = FALSE)
# plotPCA(vsd, intgroup = "condition")
# ggsave("PCA.png")


cds=estimateDispersions(dds)
vsd=varianceStabilizingTransformation(cds)
plotPCA(vsd,intgroup=c("condition"))

#How to get PCA scree plot?

## calculate the variance for each gene
rv <- rowVars(assay(vsd))

## select the ntop genes by variance
select <- order(rv, decreasing=TRUE)[seq_len(min(500, length(rv)))]

## perform a PCA on the data in assay(x) for the selected genes
pca <- prcomp(t(assay(vsd)[select,]))

## the contribution to the total variance for each component
percentVar <- (pca$sdev^2 / sum( pca$sdev^2 ))*100

##plot the "percentVar"
scree_plot=data.frame(percentVar)
scree_plot[,2] <- c(1:ncol(dds))
colnames(scree_plot)<- c("variance", "component_number")
ggplot(scree_plot, mapping=aes(x=component_number, y=variance))+geom_bar(stat="identity")

```

# Volcano Plot Function
```{r} 

make_volcano_pdf <- function(res_obj,
                             title_text,
                             outfile,
                             sig_cutoff = 0.05,
                             lfc_cutoff = 2,
                             top_n_up = 10,
                             top_n_down = 10) {

  df <- as.data.frame(res_obj)
  df$ensembl_gene <- rownames(df)

  df <- df[!is.na(df$log2FoldChange) & !is.na(df$pvalue), ]
  df$neglog10_p <- -log10(df$pvalue)

  # Ensembl -> symbol
  df$ensembl_clean <- sub("\\..*$", "", df$ensembl_gene)
  df$symbol <- mapIds(
    org.Hs.eg.db,
    keys = df$ensembl_clean,
    column = "SYMBOL",
    keytype = "ENSEMBL",
    multiVals = "first"
  )
  df$plot_label <- ifelse(is.na(df$symbol) | df$symbol == "", df$ensembl_clean, df$symbol)

  # categories
  df$group <- "Not significant"
  df$group[!is.na(df$pvalue) & df$pvalue < sig_cutoff & df$log2FoldChange >= lfc_cutoff] <- "Upregulated"
  df$group[!is.na(df$pvalue) & df$pvalue < sig_cutoff & df$log2FoldChange <= -lfc_cutoff] <- "Downregulated"

  df$group <- factor(df$group, levels = c("Not significant", "Upregulated", "Downregulated"))

  # labels only from significant genes that also pass fold-change cutoff
  label_pool <- df[df$group %in% c("Upregulated", "Downregulated"), ]

  up_df <- label_pool[label_pool$group == "Upregulated", ]
  up_df <- up_df[order(up_df$pvalue, -abs(up_df$log2FoldChange)), ]
  up_df <- head(up_df, top_n_up)

  down_df <- label_pool[label_pool$group == "Downregulated", ]
  down_df <- down_df[order(down_df$pvalue, -abs(down_df$log2FoldChange)), ]
  down_df <- head(down_df, top_n_down)

  lab_df <- rbind(up_df, down_df)
  lab_df <- lab_df[!duplicated(lab_df$plot_label), ]

  p <- ggplot(df, aes(x = log2FoldChange, y = neglog10_p)) +
    geom_hline(
      yintercept = -log10(sig_cutoff),
      linetype = "dashed",
      linewidth = 0.4,
      color = "grey40"
    ) +
    geom_vline(
      xintercept = c(-lfc_cutoff, lfc_cutoff),
      linetype = "dashed",
      linewidth = 0.4,
      color = "grey40"
    ) +
    geom_point(
      data = df[df$group == "Not significant", ],
      shape = 1,
      color = "black",
      size = 1.2,
      alpha = 0.8
    ) +
    geom_point(
      data = df[df$group == "Upregulated", ],
      color = "#ff0000",
      size = 1.2,
      shape = 1,
      alpha = 0.9
    ) +

    geom_point(
      data = df[df$group == "Downregulated", ],
      color = "#1f78b4",
      size = 1.2,
      shape = 1,
      alpha = 0.9
    ) +
    geom_text_repel(
      data = lab_df,
      aes(label = plot_label),
      size = 2.4,
      box.padding = 0.4,
      point.padding = 0.2,
      min.segment.length = 0,
      force = 10
    ) +
    labs(
      title = title_text,
      x = expression(log[2](fold~change)),
      y = expression(-log[10](italic(P)))
    ) +
    theme_bw() +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 11),
      axis.title = element_text(face = "bold"),
      axis.text = element_text(color = "black")
    )

  ggsave(outfile)
}
```

Making volcano plots
```{r}
make_volcano_pdf(
  all_genes_low,
  "Differential Expression in Low Mitochondrial State versus Control",
  "volcano_mitoLow_vs_control.png",
  sig_cutoff = 0.05,
  lfc_cutoff = 2
)

make_volcano_pdf(
  all_genes_high,
  "Differential Expression in High Mitochondrial State versus Control",
  "volcano_mitoHigh_vs_control.png",
  sig_cutoff = 0.05,
  lfc_cutoff = 2
)

make_volcano_pdf(
  all_genes_hl,
  "Differential Expression in High versus Low Mitochondrial State",
  "volcano_mitoHigh_vs_mitoLow.png",
  sig_cutoff = 0.05,
  lfc_cutoff = 2
)
```

Reading-in Upregulated and Downregulated Genes
```{r}
# up_high <- as.data.frame(read.csv(paste0("sig_genes_high", "_upregulated.csv")))
# down_high <- as.data.frame(read.csv(paste0("sig_genes_high", "_downregulated.csv")))
# 
# up_low <- as.data.frame(read.csv(paste0("sig_genes_low", "_upregulated.csv")))
# down_low <- as.data.frame(read.csv(paste0("sig_genes_low", "_downregulated.csv")))
# 
# up_hl <- as.data.frame(read.csv(paste0("sig_genes_hl", "_upregulated.csv")))
# down_hl <- as.data.frame(read.csv(paste0("sig_genes_hl", "_downregulated.csv")))



symbol_ids <- bitr(sig_genes_hl$gene,
                 fromType = "ENSEMBL",
                 toType = "SYMBOL",
                 OrgDb = org.Hs.eg.db)

sting_genes <- c(
  "TMEM173","TBK1","IRF3","IFNB1","CXCL10","CCL5",
  "ISG15","IFIT1","IFIT2","MX1","MX2","OAS1","OAS2","IRF7"
)

symbol_ids %>%
  filter(SYMBOL %in% sting_genes)


ifn_genes <- c(
  "IFNB1","CXCL10","CCL5","ISG15","IFIT1","IFIT2","IFIT3",
  "MX1","MX2","OAS1","OAS2","OAS3","IFI6","IFI27","IFI44",
  "IFI44L","IRF7","STAT1","RSAD2"
)

symbol_ids %>%
  filter(SYMBOL %in% ifn_genes)
```

GO Function (includes bar plot and manhattan plot)
```{r}
run_go <- function(res_obj) {
  
  res_df <- as.data.frame(res_obj)

  go_res <- enrichGO(
    gene = unique(res_obj$gene),
    keyType = "ENSEMBL",
    OrgDb = org.Hs.eg.db,
    ont = "ALL",
    pvalueCutoff = 0.05,
  )
  
  go_res@result$neglogpadj <- -log10(go_res@result$p.adjust)

  p <- barplot(
    go_res,
    x = "neglogpadj",
    showCategory = 10,
    split = "ONTOLOGY",
    label_format = 50
  ) +
    aes(fill = ONTOLOGY) +
    scale_fill_brewer(palette = "Set2") +
    enrichplot::autofacet(by = "row", scales = "free")


  ggsave(paste0(deparse(substitute(res_obj)), "_GO_barplot.png"))
  
  return(go_res)
}
```


Running GO Function
```{r}
# up_high_go_res <- run_go(up_high)
# down_high_go_res <- run_go(down_high)
# 
# up_low_go_res <- run_go(up_low)
# down_low_go_res <- run_go(down_low)
# 
# up_hl_go_res <- run_go(up_hl)
# down_hl_go_res <- run_go(down_hl)

high_go_res <- run_go(sig_genes_high)
low_go_res <- run_go(sig_genes_low)
hl_go_res <- run_go(sig_genes_hl)

high_go_res_simplified <- simplify(high_go_res, cutoff=0.7, by="p.adjust", select_fun=min)
low_go_res_simplified <- simplify(low_go_res, cutoff=0.7, by="p.adjust", select_fun=min)
hl_go_res_simplified <- simplify(hl_go_res, cutoff=0.7, by="p.adjust", select_fun=min)

```

Calculating similarity matrix and reducing GO terms
```{r}
# library(rrvgo)
# 
# go_analysis <- high_go_res
# 
# simMatrix <- calculateSimMatrix(go_analysis$ID,
#                                 orgdb="org.Hs.eg.db",
#                                 ont="BP",
#                                 method="Rel")
# 
# scores <- setNames(-log10(go_analysis$qvalue), go_analysis$ID)
# reducedTerms <- reduceSimMatrix(simMatrix,
#                                 scores,
#                                 threshold=0.7,
#                                 orgdb="org.Hs.eg.db")
# 
# heatmapPlot(simMatrix,
#             reducedTerms,
#             annotateParent=TRUE,
#             annotationLabel="parentTerm",
#             fontsize=6)
# 
# scatterPlot(simMatrix, reducedTerms)
# 
# treemapPlot(reducedTerms)
```

Calculating similarity matrix and reducing GO terms
```{r}
# go_analysis <- hl_go_res
# 
# simMatrix <- calculateSimMatrix(go_analysis$ID,
#                                 orgdb="org.Hs.eg.db",
#                                 ont="BP",
#                                 method="Rel")
# 
# scores <- setNames(-log10(go_analysis$pvalue), go_analysis$ID)
# reducedTerms <- reduceSimMatrix(simMatrix,
#                                 scores,
#                                 threshold=0.7,
#                                 orgdb="org.Hs.eg.db")
# 
# 
# scatterPlot(simMatrix, reducedTerms)
# 
# treemapPlot(reducedTerms)
```

Calculating similarity matrix and reducing GO terms
```{r}
# go_analysis <- low_go_res
# 
# simMatrix_low <- calculateSimMatrix(go_analysis$ID,
#                                 orgdb="org.Hs.eg.db",
#                                 ont="BP",
#                                 method="Rel")
# 
# scores_low <- setNames(-log10(go_analysis$pvalue), go_analysis$ID)
# reducedTerms_low <- reduceSimMatrix(simMatrix_low,
#                                 scores_low,
#                                 threshold=0.7,
#                                 orgdb="org.Hs.eg.db")
# 
# 
# scatterPlot(simMatrix_low, reducedTerms_low)
# 
# treemapPlot(reducedTerms_low)
```



KEGG Function
```{r}
run_kegg <- function(res_obj) {
  
  res_df <- as.data.frame(res_obj)

  entrez_ids <- bitr(res_df$gene,
                 fromType = "ENSEMBL",
                 toType = "ENTREZID",
                 OrgDb = org.Hs.eg.db)

  kegg_res <- enrichKEGG(
    gene = unique(entrez_ids$ENTREZID),
    organism = "hsa",
    pvalueCutoff = 0.05
  )
  
  # Makes a dot plot of the results
  dotplot(kegg_res, showCategory = 15)
  ggsave(paste0(deparse(substitute(res_obj)), "_KEGG_dotplot.png"))
  
  return(kegg_res)
}
```

Running KEGG
```{r}
# up_high_kegg_res <- run_kegg(up_high)
# down_high_kegg_res <- run_kegg(down_high)
# 
# up_low_kegg_res <- run_kegg(up_low)
# # down_low_kegg_res <- run_kegg(down_low)# only one gene
# 
# up_hl_kegg_res <- run_kegg(up_hl) 
# down_hl_kegg_res <- run_kegg(down_hl)
# 
# nrow(up_high_kegg_res)
# nrow(down_high_kegg_res)
# nrow(up_low_kegg_res)
# nrow(up_hl_kegg_res)
# nrow(down_hl_kegg_res)

high_kegg_res <- run_kegg(sig_genes_high)
low_kegg_res <- run_kegg(sig_genes_low)
hl_kegg_res <- run_kegg(sig_genes_hl)


```

```{r}
library(ggtree)

edox2 <- pairwise_termsim(high_go_res)
edox3 <- pairwise_termsim(low_go_res)
edox4 <- pairwise_termsim(hl_go_res)

treeplot(edox2)
treeplot(edox3)
treeplot(edox4)

ssplot(edox2, group_category = FALSE, node_label = "group")
ggsave("high_go_ssplot.png")
ssplot(edox3, group_category = FALSE, node_label = "group")
ggsave("low_go_ssplot.png")
ssplot(edox4, group_category = FALSE, node_label = "group")
ggsave("hl_go_ssplot.png")
```


